---
title: "Workflow for Risa"
date: '`r format(Sys.Date())`'
output:
  html_document:
    toc: true
    toc_float: true
    dev: 'svg'
    number_sections: true
    pandoc_args: --lua-filter=color-text.lua
    highlight: pygments 
    theme: 'flatly'
---

# Notes

1. **Great resources for learning R**:
    + *R for Data Science*, Hadley Wickham and Garrett Grolemund: <https://r4ds.had.co.nz/>
    + *Happy Git and GitHub for the useR*, Jenny Bryan and Jim Hester: <https://happygitwithr.com/>
  
1. This workflow generally follows Ilastik's animal tracking workflow described here: <https://www.ilastik.org/documentation/animaltracking/animaltracking>


# Load `R` libraries

In `R`:

```{r, message =F, warning=F}
library(tidyverse)
library(ijtiff)
library(magick)
```

## External software 

You'll also need the following external software:

* Fiji: <https://imagej.net/Downloads>
* Ilastik: <https://www.ilastik.org/download.html>

# Workflow

## Create file with list of samples

You have the metadata file called `samples.csv`. Here you want to use it to create a file with just a list of samples, so that you can "loop", i.e. successively run the following commands, over each of the samples.

In `bash`:

```{bash, eval =F}
# NOTE: $samples_metadata_file is the file I sent called `samples.csv`
################################################## To adapt
samples_metadata_file=/Users/risa_admin/Desktop/test/samples.csv
samples_list=/User/risa_admin/Desktop/test/samples_list.csv
##################################################

# Copy first column, remove header, and send to new file containing a list of samples to analyse
cut -f1 -d',' $samples_metadata_file | tail -n+2 > $samples_list
```

## Copy videos from FTP to working directory

Copy over the videos to your working directory.

In `bash`: 

```{bash, eval = F}
################################################## To adapt
path_to_ftp={path_to_ftp_directory}
path_to_wd={path_to_working_directory}
################################################## 

cp -r $path_to_ftp/* path_to_wd
```

## Draw lanes

This is useful because at the end of the Ilastik workflow, the software will output a CSV with the coordinates of all objects it thinks are fish across all the frames in each video. However, if it loses track of a fish in some frames and then picks it up again later, it will think that it's a different fish. So in your analysis you need to assume that any object identified within a given lane (i.e. within the defined interval of y-coordinates) is *the same* fish, by reassigning the labels. We'll return to this issue after the tracking is complete.

The following code extracts a single frame from the video, draws horizontal lines where you've specified the boundaries of the lanes, and then saves it to a file. If the lane coordinates needs to be adjusted, you can just change the numbers in the appropriate columns in the `samples.csv` file. 

In `R`:

```{r, eval = F}
################################################## To adapt
samples_meta_file="/Users/risa_admin/Desktop/test/samples.csv"
video_dir="/Users/risa_admin/Desktop/test/videos"
output_path="/Users/risa_admin/Desktop/test/lanes"
################################################## 

# Read in samples file
samples_df = readr::read_csv(samples_meta_file)

# Loop over each video
lapply(unique(samples_df$SAMPLE), function(sample_name){
  
  sample_video = file.path(video_dir, paste(sample_name, ".ome.tif", sep = ""))
  
  # Get lane coords
  
  ## Total lanes in video
  lanes_n = samples_df %>%
    dplyr::filter(SAMPLE == sample_name) %>%
    dplyr::pull(TOTAL_LANES)
  
  ## Get lane coords
  lane_coords = samples_df %>%
    dplyr::filter(SAMPLE == sample_name) %>%
    dplyr::select(starts_with("END_LANE")) %>%
    subset(select = 1:lanes_n -1) %>%
    unlist(use.names = F)
  
  # read first frame of tiff
  frame_1 = ijtiff::read_tif(sample_video, frames = 1)
  
  # write to file (magick can't read directly...)
  tmp_file_name = paste(sample_name, ".tmp.tif", sep = "")
  ijtiff::write_tif(frame_1, tmp_file_name)
  
  # read back in as magick image
  frame_1_m = magick::image_read(tmp_file_name)
  
  # add horizontal lines
  
  img = magick::image_draw(frame_1_m) # make image object
  abline(h = lane_coords, col = "white") # draw lines
  lined_img = magick::image_scale(img, 1000) %>% # shrink
      magick::image_modulate(brightness = 700) # increase brightness
  dev.off()
  
  # write image
  
  ## Create output name
  out_name = file.path(output_path, paste(sample_name, ".png"))
  magick::image_write(lined_img, path = out_name, format = "png")
  
  # clean up
  
  file.remove(tmp_file_name)
  
})
```

## Extract training frames

The Ilastik GUI can struggle when dealing with large videos. To keep it relatively fast, you want to create short subsets of the videos and lead them into Ilastik for training instead.

In `R`:

```{r, eval = F}
################################################## To adapt
path_to_samples_file="/Users/risa_admin/Desktop/test/samples.csv"
sample_video="{path_to_sample_video_tif_to_analyse}"
training_video_dir="{path_to_output_directory}"
################################################## 

# Read in samples file
samples_df = readr::read_csv(path_to_samples_file)

# Loop over each video
lapply(unique(samples_df$SAMPLE), function(sample_name){
  
  sample_video = file.path(video_dir, paste(sample_name, ".ome.tif", sep = ""))

  # Get start and end frames
  start = samples_df %>% 
    dplyr::filter(SAMPLE == sample_name) %>% 
    dplyr::pull(SAMPLE_START)
  
  end = samples_df %>% 
    dplyr::filter(SAMPLE == sample_name) %>% 
    dplyr::pull(SAMPLE_END)
  
  # Extract training frames
  tif = ijtiff::read_tif(sample_video,
                         frames = start:end)

  # Write file
  ## Create output name
  out_file = file.path(training_video_dir, paste(sample_name, ".tif"))
  ijtiff::write_tif(tif, out_file)

})
```

## Convert tifs to h5

Ilastik requires videos in the `.h5` format. To convert the videos from their original `.tif` format to `.h5`, you can use the Ilastik plugin in Fiji. 

In `bash`:

```{bash, eval = F}
## NOTE: run this on both short training video tif AND full video tif by changing $input_directory
################################################## To adapt
samples_list=/Users/risa_admin/Desktop/test/samples.csv
convert_tif2h5_script=/Users/risa_admin/Desktop/test/convert_tif_to_h5.ijm
fiji={path_to_fiji}
input_directory={tif_input_path}
output_directory={h5_output_path}
################################################## 

# Loop over each video in $samples_list
for sample in $(cat $samples_list ); do \
  # Create $input_path and $output_path (note  that Fiji requires the absolute (not relative) paths of files, hence `realpath`)
  input_file=$(realpath $input_directory/$sample.ome.tif) ;
  output_file=$(realpath $output_directory/$sample.h5) ;
  # Create string of parameters
  params=$(echo -e \'input_tif\=\"$input_file\",output_h5\=\"$output_file\"\') ;
  # Create command
  cmd=$(echo "$fiji" --ij2 --headless --console --run "$convert_tif2h5_script" "$params") ;
  # Run Fiji
  eval $cmd ;
done
```

## Create Ilastik pixel classification project and train on training videos

Create a pixel classification project as described here: <https://www.ilastik.org/documentation/animaltracking/animaltracking#pixel-foregroundbackground-segmentation-using-the-pixel-classification-workflow>.

Then train the project with your shortened training videos.

## Generate `{sample}_Probabilities.h5` files with trained Ilastik project

Run the trained project on all training **AND** full videos to create the `{sample}_Probabilities.h5` files.

In `bash`:

```{bash, eval = F}
# Warning: this code works with Ilastik v1.4.0b13. Other versions may throw errors.
##################################################
samples_list={path_to_samples_list}
ilastik_run={path_to_ilastik_runscript} # e.g. /ilastik-release/run_ilastik.sh
pixclass_proj={path_to_ilastik_pixel_classification_project}
input_directory={h5_input_directory}
##################################################

for sample in $(cat $samples_list); do \
  # Set input video filename and output filename
  input=$input_directory/$sample.h5 ;
  output=$input_directory/$sample\_Probabilities.h5 ;
  # Run Ilastik
  $ilastik_run \
    --headless \
    --project=$pixclass_proj \
    --readonly \
    --raw_data=$input \
    --export_source="Probabilities" \
    --output_filename_format=$output ;
done
```

## Create Ilastik animal tracking project and train on training videos

Create an Ilastik animal tracking project as described here: <https://www.ilastik.org/documentation/animaltracking/animaltracking#automatic-tracking-using-the-animal-tracking-workflow>.

Then load up the training videos **AND** their respective `{sample}_Probabilities.h5` files, and train the project as described in the workflow.

Once it's trained, you can run it on all the full videos to create the CSV output:

In `bash`:

```{bash, eval = F}
##################################################
samples_list={path_to_samples_list}
ilastik_run={path_to_ilastik_runscript} # e.g. /ilastik-release/run_ilastik.sh
anitrack_proj={path_to_ilastik_animal_tracking_project}
input_directory={path_to_video_input_directory}
output_directory={path_to_csv_raw_output_directory}
##################################################

for sample in $(cat $samples_list); do \
  # Set input video and probabilities filenames, and output filename
  input_vid=$input_directory/$sample.h5 ;
  input_probs=$input_directory/$sample\_Probabilities.h5 ;
  output_csv=$output_directory/$sample\_CSV-Table.csv ;
  # Run Ilastik
  $ilastik_run \
    --headless \
    --readonly \
    --project=$anitrack_proj \
    --raw_data=$input_vid \
    --input_axes=\"tzyxc\" \
    --prediction_maps=$input_probs \
    --export_source=\"Plugin\" \
    --export_plugin=\"CSV-Table\" \
    --output_filename_format=$output_csv ;
done

# NOTE: for some reason, Ilastik may output the file with an extra `.csv` in the filename
# e.g. $output_directory/$sample_CSV-Table.csv.csv
```

## Extract key columns from CSV

The CSV output contains a lot of data that you don't need, so this step extracts the columns you do need.

In `R`: 

```{r, eval = F}
##################################################
raw_csv_dir = "{path_to_csv_raw_output_directory}"
new_csv_dir = {path_to_cleaned_csv_directory}
##################################################

files_to_process = list.files(raw_csv_dir, full.names = T)

lapply(files_to_process, function(FILE){
  # Get sample name
  SAMPLE = basename(FILE) %>% 
    stringr::str_remove("_CSV-Table.csv.csv")
  # Create output file name
  out_file = file.path(new_csv_dir, paste(SAMPLE, ".csv", sep = ""))
  # Read in file, extract key columns, and write to new directory
  readr::read_csv(FILE) %>%
    dplyr::select(FRAME = frame,
                  TRACKID = trackId,
                  COORD_X = Object_Center_0,
                  COORD_Y = Object_Center_1) %>%
    readr::write_csv(out_file)  
})
```

## Read CSVs into R and create tracking plots

### Read in files

In `R`:

```{r, eval = F}
##################################################
samples_file = "{path_to_samples_metadata_file}"
in_dir = "{path_to_cleaned_csvs_directory}"
##################################################

# Get samples metadata
samples_df = readr::read_csv(samples_file)

# Get list of files
files = list.files(in_dir, full.names = T)
names(files) = basename(files) %>% 
  stringr::str_remove(".csv")

# Read all into list
df_list = lapply(files, function(file){
  # get sample name
  sample_name = basename(file) %>% 
    stringr::str_remove(".csv")
  # get target_row from samples_df
  target_row = samples_df %>% 
    dplyr::filter(SAMPLE == sample_name)
  # set up output list
  out = list()
  # get sample name
  out[["META"]] = list()
  out[["META"]][["SAMPLE"]] = sample_name
  # add some metadata
  out[["META"]][["TOTAL_FRAMES"]] = target_row %>% 
    dplyr::pull(TOTAL_FRAMES)
  out[["META"]][["MAX_Y"]] = target_row %>% 
    dplyr::pull(MAX_Y)
  out[["META"]][["N_LANES"]] = target_row %>% 
    dplyr::pull(TOTAL_LANES)
  
  # add lane breaks
  lanes_n = out$META$N_LANES
  
  out[["META"]][["LANE_BREAKS"]] = samples_df %>%
    dplyr::filter(SAMPLE == out$META$SAMPLE) %>%
    dplyr::select(starts_with("END_LANE")) %>%
    subset(select = 1:lanes_n -1) %>%
    unlist(use.names = F) %>% 
    c(0, ., out$META$MAX_Y)
  
  # add tracking data
  out[["RAW"]] = readr::read_csv(file)
  
  # process
  ## create recode vector for `COORD_Y`
  frames_seq = seq(0, out$META$MAX_Y)
  recode_vec = rev(frames_seq)
  names(recode_vec) = frames_seq
  ## run
  out[["CLEAN"]] = out$RAW %>% 
    dplyr::filter(TRACKID != -1) %>% 
    # divide by lane and flip Y coords
    dplyr::mutate(LANE = cut(COORD_Y,
                             breaks = out$META$LANE_BREAKS,
                             labels = F),
                  COORD_X = round(COORD_X),
                  COORD_Y = round(COORD_Y),
                  COORD_Y = dplyr::recode(COORD_Y, !!!recode_vec)) %>% 
    dplyr::arrange(LANE, FRAME) 
  
  return(out)
})

return(df_list)

```

### Create tile plot

This quality-control step creates a tile plot. It fills a tile for each frame in each video where an object is tracked, so you can see where the tracking failed, or where the fish was hidden.

In `R`: 

```{r, eval = F}
##################################################
out_path = "{directory_for_tile_plots}"
##################################################

# create directory if it doesn't exist
dir.create(out_path, showWarnings = F, recursive = T)
out_file = file.path(out_path, "tile.png")

# extract data from list and bind into DF
lapply(df_list, function(SAMPLE){
  SAMPLE$CLEAN
}) %>% 
  dplyr::bind_rows(.id = "SAMPLE") %>% 
  dplyr::mutate(LANE = factor(LANE, levels = rev(1:max(LANE)))) %>% 
  ggplot() +
    geom_tile(aes(FRAME, LANE, fill = LANE)) +
    facet_wrap(~SAMPLE, ncol = 1) +
    scale_fill_viridis_d() +
    guides(fill = guide_legend(reverse = T))

# save
ggsave(out_file, 
       device = "png",
       width = 20,
       height = 40,
       units = "cm",
       dpi = 400)  
```

### Create path plot

The path plot creates lines that follows the position of each fish over the course of the video.

In `R`: 

```{r, eval = F}
##################################################
out_path = "{directory_for_path_plots}"
##################################################

dir.create(out_path, recursive = T)
out_file = file.path(out_path, "path.png")

lapply(df_list, function(SAMPLE){
  SAMPLE$CLEAN
}) %>% 
  dplyr::bind_rows(.id = "SAMPLE") %>% 
  ggplot() +
  geom_path(aes(COORD_X, COORD_Y, group = LANE, colour = FRAME)) +
  scale_colour_viridis_c() +
  coord_fixed() +
  facet_wrap(~SAMPLE, ncol = 1) +
  theme(strip.text = element_text(size = 8)) +
  # save
  ggsave(out_file, 
         device = "png",
         width = 30,
         height = 100,
         units = "cm",
         dpi = 400)
```

### Collapse all CSVs into single CSV and write to table

In `R`: 

```{r, eval = F}
##################################################
out_path = "{directory_for_final_consolidated_CSV}"
##################################################

dir.create(out_path, recursive = T)
out_file = file.path(out_path, "final.csv")

lapply(df_list, function(SAMPLE){
  SAMPLE$CLEAN
}) %>%
  dplyr::bind_rows(.id = "SAMPLE") %>%
  dplyr::select(SAMPLE, LANE, FRAME, COORD_X, COORD_Y) %>% 
  readr::write_csv(out_file)
```


